Light inextensible strings under tension in the Schwarzschild geometry

### 2019-07-11 string_grg.tex
multiply defined labels
more work

\section{Introduction}
\label{intro}

\section{Introduction} % Section titles are automatically converted to all-caps.
% Section numbering is automatic.

\label{intro}
Light inextensible strings are a stalwart feature of elementary
physics.  A light string is not subject to gravity, and being under
tension means that its configuration is an extremum of proper length.
Of course, no real string is perfectly weightless or inextensible.
With tension~$T$, gravitational acceleration $g$ and (proper) mass per
(proper) unit length~$\mu$, then a string will behave as though light
on lengthscales $L\lesssim T/\left(\mu g\right)$.  Using Newtonian
expressions for the strength of gravity at the event horizon of a
black hole with mass $M$ gives $L\lesssim TM/\left(\mu c^4\right)$.
For the string to behave as though light on lengthscales comparable
with the Schwarzschild radius would require $T\gtrsim\mu c^2$.
Exceeding this criterion would violate the dominant energy condition
(and in this context we note that transverse waves would propagate at
speed~$\sqrt{T/\mu}\sim c$).  Nevertheless, the idealization of a
string as light and inextensible is an interesting and informative
limit: its analysis interprets the concept ``straight line'' and
suggests a possible way of manipulating a black hole.  Peculiar
counterintuitive behaviour such as inward-directed centrifugal force
in the vicinity of a black hole is well
known~\cite{abramowicz1992,abramowicz1993}, and the analysis of taut
strings furnishes another everyday window into general relativity.
Another possibility, discussed by Thorne~\cite{thorne1986}, is to pass
electric current through such taut strings, an option not pursued
here.

>>>>

\section{Introduction}

Light inextensible strings are a stalwart feature of elementary
physics.  A light string is not subject to gravity, and being under
tension means that its configuration is an extremum of proper length.
Of course, no real string is perfectly weightless or inextensible.
With tension~$T$, gravitational acceleration $g$ and (proper) mass per
(proper) unit length~$\mu$, then a string will behave as though light
on lengthscales $L\lesssim T/\left(\mu g\right)$.  Using
$c^4/\left(4GM\right)$ for the strength of gravity at the event
horizon of a black hole with mass $M$ gives $L\lesssim TM/\left(\mu
c^4\right)$.  For the string to behave as though light on lengthscales
comparable with the Schwarzschild radius would require $T\gtrsim\mu
c^2$.  Exceeding this criterion would violate the dominant energy
condition (and in this context we note that transverse waves would
propagate at speed~$\sqrt{T/\mu}\sim c$).  Nevertheless, the
idealization of a string as light and inextensible is an interesting
and informative limit: Rindler~\cite{rindler}, for example, considers
lowering a heavy object to the event horizon on the end of such a
string to illustrate energetic arguments in GR.

Analysis of non-radial strings interprets the concept ``straight
line'' and suggests a possible way of manipulating a black hole.
Peculiar counterintuitive behaviour such as inward-directed
centrifugal force in the vicinity of a black hole is well
known~\cite{abramowicz1992,abramowicz1993}, and the analysis of taut
strings furnishes another everyday window into general relativity.
Another possibility, discussed by Thorne~\cite{thorne1986}, is to pass
electric current through such taut strings, an option not pursued
here.
